---
title: "Bitcoin Prices: A Real-World Example"
---

This example applies change-point detection to hourly Bitcoin closing prices.
The data are derived from the
[Bitcoin Historical Data](https://www.kaggle.com/datasets/mczielinski/bitcoin-historical-data)
dataset on Kaggle, which contains minute-level OHLC prices and trading volume.

## Load the Data

```{r}
btc_hourly <- read.csv(
  "https://raw.githubusercontent.com/Prabashoka/scanr-vignette/refs/heads/main/Data/BTC-hourly-data-closing-price.csv"
)

btc_hourly$Datetime <- as.POSIXct(
  btc_hourly$Datetime,
  tz = "UTC"
)

head(btc_hourly)
```
Load the installed `scanr` package.
```{r}
library(scanr)
```

## Choose Window Sizes

Following the Python analysis, the largest candidate window is
$\lfloor n^{2/3}\rfloor$. The package helper selects 15 default window sizes
between 100 hours and that upper bound.

```{r}
n <- nrow(btc_hourly)
window_sizes <- default_window_sizes(
  n,
  min_window = 100,
  max_window = floor(n^(2 / 3)),
  n_windows = 15,
  seed = 1000
)
window_sizes
```

The closing prices are standardized before detection. Standardization changes
their units but preserves the locations of mean shifts.

```{r}
x <- as.numeric(scale(btc_hourly$Close))
```

## Detect Changes in Mean

The Python voting threshold of 0.625 corresponds to
`vote_threshold = 0.625` in `scan_cpd()`. The R interface uses `n_boot` and
`random_state` for bootstrap calibration and reproducibility.

```{r}
fit_btc <- scan_cpd(
  x,
  window_sizes = window_sizes,
  n_boot = 1000,
  vote_threshold = 0.5,
  random_state = 100,
  change_type = "mean",
  n_jobs = 1
)

fit_btc
```

The fitted object stores change points as positions in the hourly series. We
can map those positions back to UTC dates and observed prices.

```{r}
detected_changes <- btc_hourly[
  fit_btc$change_points,
  c("Datetime", "Close"),
  drop = FALSE
]

cat("Number of change-points:", nrow(detected_changes), "\n")
detected_changes
```

## Visualize the Result

```{r}
#| fig-width: 11
#| fig-height: 5

vis_change_points(
  x,
  fit_btc,
  index = btc_hourly$Datetime,
  x_label = "Date (UTC)",
  y_label = "Standardized hourly closing price",
  title = "Mean changes in hourly Bitcoin closing prices"
)
```

## Inspect the Vote Scree

The vote scree shows how much support each candidate location receives across
the selected window sizes. The horizontal threshold corresponds to the
`vote_threshold = 0.5` used above.

```{r}
#| fig-width: 11
#| fig-height: 5

vis_vote_scree(fit_btc)
```

::: {.callout-caution}
Bitcoin prices are strongly trending and serially dependent. Detected changes
identify shifts in the mean price level; they should not be interpreted as
causal market events.
:::
