# Change in Mean (Example 1)

```{r}
#| label: setup-mean
#| include: false

library(scanr)
```

We first simulate a white-noise time series of length 20,000 with 20 change
points and 21 piecewise-constant mean regimes. The noise variance is fixed, so
this example focuses on changes in the mean.

```{r}
set.seed(1234)

n <- 20000

change_points <- c(
  952, 1905, 2858, 3810, 4763, 5715, 6668, 7620, 8573, 9525,
  10478, 11430, 12383, 13335, 14288, 15240, 16193, 17145, 18098,
  19050
)

means <- c(
  0, 2, -1, 3, 0.5, -2, 2, 5, -0.5, 2.5, 0, -2.5, -1.5, 1.5,
  3, 1, 0, 1.25, -2, 3.5, -1.5
)

segment_starts <- c(1L, change_points + 1L)
segment_ends <- c(change_points, n)
x_mean <- numeric(n)

for (j in seq_along(means)) {
  segment_index <- segment_starts[j]:segment_ends[j]
  x_mean[segment_index] <- rnorm(length(segment_index), mean = means[j], sd = 1)
}

change_points
```

The plot below shows the simulated series.

```{r}
#| echo: false
#| warning: false
#| fig.width: 10
#| fig.height: 4

vis_time_series(
  x_mean,
  x_label = "Time",
  y_label = "Value",
  title = ""
)
```

## Run SCAN Change-Point Detection

The `window_sizes` argument controls the local scales used by the scan. The
window sizes should be smaller than the spacing between nearby change points,
while still being large enough to estimate the local distributions on each side
of a candidate split.

```{r}
fit_mean <- scan_cpd(
  x_mean,
  window_sizes = c(100, 164, 227, 291, 355, 418, 482, 546, 609, 673, 737),
  n_boot = 400,
  random_state = 1234,
  change_type = "mean",
  n_jobs = 1
)

fit_mean
```

::: {.callout-note}
## Remark on Bootstrap Calibration

The adaptive threshold for declaring a change point is derived through a
**tapered block bootstrap** across the combined windows. Increasing the number
of bootstrap samples (`n_boot`) produces a more accurate null distribution at
the cost of longer computation time; a value of 400-1,000 is typically
sufficient. The method supports parallel processing via
[Rust's](https://www.rust-lang.org/) **Rayon** library for computational
efficiency.
:::

The estimated change points can be extracted directly from the fitted object.

```{r}
fit_mean$change_points
```

## Evaluate Detection Accuracy

For simulated data, estimated change points can be compared with the known
truth. The `tolerance` argument controls how close an estimated change point
must be to a true change point to count as a match.

```{r}
cpd_metrics(
  true_cps = change_points,
  estimated_cps = fit_mean$change_points,
  n = length(x_mean),
  tolerance = 20
)
```

## Visualize the Result

The package includes visual helpers for inspecting detected change points,
window-level votes, and the overall vote scree.

```{r}
#| fig.width: 10
#| fig.height: 4.5

vis_change_points(
  x_mean,
  fit_mean,
  true_change_points = change_points,
  x_label = "Time",
  y_label = "Value"
)
```

## Tune the Voting Threshold

The ensemble vote threshold, `vote_threshold`, controls how much agreement is
required across window sizes before a candidate is retained. The default value
is 0.5, meaning that a candidate must be supported by at least half of the
available window-level votes.

```{r}
vis_vote_scree(fit_mean)
```

```{r}
#| fig.width: 10
#| fig.height: 4.5

vis_window_votes(fit_mean)
```
