# SWAL Statistic for Single Change-Point Detection

```{r}
#| label: setup-swal
#| include: false

library(Scanr)
```

The SCAN procedure above combines local evidence across many windows. The
package also exposes lower-level localization tools for a single region that is
believed to contain one change point.

For a mean change, the mean version of the SWAL statistic aligns with the CUSUM
localizer. Both methods search for the split that best separates the left and
right sides of the local region.

```{r}
set.seed(1234)

true_single_cp <- 150
mean_region <- c(
  rnorm(true_single_cp, mean = 0, sd = 1),
  rnorm(true_single_cp, mean = 2, sd = 1)
)

c(
  truth = true_single_cp,
  cusum = ts_cusum(mean_region),
  swal_distribution = swal_statistic(mean_region, change_type = "distribution")
)
```

The distributional version is more general. It can localize changes where the
mean is approximately unchanged but other aspects of the distribution, such as
variance, change sharply. In the example below, both segments have mean zero,
but the second segment has a much larger standard deviation.

```{r}
set.seed(1234)

var_region <- c(
  rnorm(true_single_cp, mean = 0, sd = 0.5),
  rnorm(true_single_cp, mean = 0, sd = 2)
)

c(
  truth = true_single_cp,
  cusum = ts_cusum(var_region),
  swal_distribution = swal_statistic(var_region, change_type = "distribution")
)
```

The `ts_wasserstein()` function returns both the estimated split and the full
sequence of split statistics. The plotting helper `vis_swal_curve()` displays
that localization curve, making it easier to inspect where the statistic is
maximized.

```{r}
vis_swal_curve(
  var_region,
  start = 1,
  end = length(var_region),
  x_label = "Candidate split",
  y_label = "Scaled Wasserstein statistic",
  title = ""
)
```
